---
title: "Argentinean League Football Analysis (2016-2022)"
sub-title: "EDA + statistical modelling"
output: github_document
---


```{r setup, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```


```{r,  warning=FALSE, message=FALSE, error=TRUE, echo=FALSE}
#----------------------------
# Import libraries and data
#----------------------------
library(ggplot2)
library(readxl)
library(dplyr)
library(readr)
library(viridis)
library(skellam)
library(formatR)

pathBase = dirname(rstudioapi::getSourceEditorContext()$path)
partidos <- read_csv(paste0(pathBase,"/partidos.csv"))

#----------------------------
#  Aesthethic variables
#----------------------------
colores = c(
    '2016-2017' ='yellow',
    '2017-2018' = "green",
    '2018-2019' = "orange",
    '2019-2020' = 'pink',
    '2021' = 'magenta',
    '2022' = "red")

fuentes_customizadas = list( theme(
                text = element_text(size=14, color = "black"),
                axis.text.x = element_text(face="plain",size = 11, color="black"),
                axis.text.y = element_text(face="plain", size = 11, color="black"),
                plot.title = element_text(size = 12, face = "bold"),
                legend.position = "none",
                legend.key.size = unit(0.5, 'cm'),
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                panel.background = element_blank(), 
                axis.line = element_line(colour = "black")), 
                labs(fill = ""))
```


# **SEASON-WISE ANALYSIS**

The first thing to evaluate is how many goals were scored per season. We can conclude from this simple analysis that the 2016-17 season had the highest number of goals scored while the 2019-20 (probably due to the suspension of lots of matches due to the COVID-19 pandemic) was the poorest one.

## 1.- Total goals per season

```{r, warning=FALSE, message=FALSE, tidy=TRUE, echo=FALSE}
#-------------------
#   Data processing
#-------------------
partidos <- na.omit(partidos)
partidos$goles_total = partidos$local_goles + partidos$visitante_goles

goles_tot = partidos %>% group_by(temporada) %>% 
    dplyr::summarise(n_goles_total = sum(goles_total))
```

```{r, echo=FALSE}
GOLES.TOTALES = ggplot(goles_tot , aes(x=temporada, y=n_goles_total, fill=temporada, color=temporada)) + 
    geom_bar(stat = "identity", width=0.8) +
    fuentes_customizadas +
    scale_fill_manual(values =colores ) +
    xlab("Season")+
    ylab("Total goals")+
    ggtitle("Total goals per season") +
    labs(fill = "")+
    theme(axis.text.x = element_text(angle = 25, hjust = 1))
```


```{r, echo=FALSE, fig.align = 'center', fig.width=6, fig.height=4}
GOLES.TOTALES
```


## 2.- Goals per season - Home vs Away

Let´s begin with the home effect analysis. The following back-to-back plot allows a comparison per season of the number of goals scored by the home and away teams. At a qualitatively level, we can see that the home team tends to score more goals that the away team. We can easily perceive there is indeed such an effect, and this type of analysis could be done match-wise within a season.

```{r,  warning=FALSE, message=FALSE, tidy=TRUE, echo=FALSE }
#-------------------
#   Data processing
#-------------------
goles_local = partidos %>% group_by(temporada) %>% 
    dplyr::summarise(n_goles_local = sum(local_goles))

goles_visitante = partidos %>% group_by(temporada) %>% 
    dplyr::summarise(n_goles_visitante = sum(visitante_goles))

goles = merge(goles_local, goles_visitante, by='temporada')
goles$temporada = as.factor(goles$temporada)

goles_long = reshape2::melt(goles, 
                            value.name = "nro_goles",
                            variable.name = "localia",
                            id.vars=c("temporada")) 
goles_df = goles_long %>% 
    mutate(nro_goles = ifelse(localia=='n_goles_visitante', -nro_goles, nro_goles)) 

temp_df <-
    goles_df %>% 
    filter(localia == "local") %>% 
    arrange(nro_goles)

the_order <- temp_df$temporada
```


```{r,  warning=FALSE, message=FALSE, tidy=TRUE, echo=FALSE }
#----------------
#     PLOT
#----------------
GOLES.LOCALIA <- ggplot(data = goles_df,
                       aes(x = temporada, y = nro_goles , group = localia, fill = localia)) +
    geom_bar(stat = "identity", width = 0.75) +
    coord_flip() +
    fuentes_customizadas +
    scale_y_continuous(breaks = seq(-600, 700, 100), 
                       labels = abs(seq(-600, 700, 100))) +
    labs(x = "Season", y = "Goals", title = "") +
        scale_fill_manual(values= c('n_goles_local'= "springgreen", 
                                    'n_goles_visitante'= "steelblue"),
                      labels = c("n_goles_local"= 'Home' , 'n_goles_visitante'='Away'),
                      name="",
                      breaks=c("n_goles_local", "n_goles_visitante")) +
    theme(legend.position = 'bottom')+
    theme(axis.text.x = element_text(angle = 25, hjust = 1))+
    ggtitle("Home and away team: back-to-back")
```


```{r, echo=FALSE, fig.align = 'center', fig.width=6, fig.height=4}
GOLES.LOCALIA  
```


## 3.- Heatmap: goals per week of the season

This heat map shows the number of goals scored by match week per season. We may get several insights from it.

* The last match week in the last two seasons showed a whooping number of goals scored in comparison to the rest of the seasons.
* Because of different formats applied to the Argentinean league, the length of the season was never kept constant. 
* The 2019-20 season was, definitively, the shortest due to the pandemic.

```{r, echo = FALSE, message=FALSE, warning=FALSE}
#----------------
#     PLOT
#----------------
heat_df = partidos %>%
    group_by(temporada, semana) %>%
    dplyr::summarise(goles = sum(goles_total))

GOLES.X.SEMANA = ggplot(heat_df, aes(x = semana, y = temporada, fill = goles)) +
    geom_tile(color = "white",lwd = 1.5, linetype = 1) + 
    coord_fixed(ratio = 3) + 
    theme_bw()+
    fuentes_customizadas+
    geom_text(aes(label =  goles), color = "white", size = 4) +
    scale_fill_gradient(low = "blue4", high = "green") +
    ylab("Season") +
    xlab("Matchweek") +
    labs(fill = "Goals") +
    ggtitle("Heat map: Number of goals per matchweek per season")
```


```{r, echo=FALSE, fig.align = 'center', fig.width=12, fig.height=8}
GOLES.X.SEMANA
```


## 4.- Heatmap: goals per match of the season

This heat map shows the average number of goals scored per match, by match week per season. With these values we can see what is the expected (I use this term as the average, indistinctly) number of goals per match.

```{r, echo = FALSE, message=FALSE, warning=FALSE}
#----------------
#     PLOT
#----------------
heat_mean = partidos %>%
    group_by(temporada, semana) %>%
    dplyr::summarise(goles_media = round(mean(goles_total),1))


GOLES.X.PARTIDO = ggplot(heat_mean, aes(x = semana, y = temporada, fill = goles_media)) +
    geom_tile(color = "white",lwd = 1.5, linetype = 1) + 
    coord_fixed(ratio = 2) + 
    theme_bw()+
    fuentes_customizadas+
    geom_text(aes(label =  goles_media), color = "white", size = 4) +
    scale_fill_gradient(low = "blue4", high = "green") +
    ylab("Season")+
    xlab("Matchweek")+
    labs(fill = "Mean goals") +
    ggtitle("Heat map: Mean number of goals per matchweek per season")
```


```{r,echo=FALSE, fig.align = 'center', fig.width=12, fig.height=8}
GOLES.X.PARTIDO
```



# **CLUB-WISE ANALYSIS**

## 1.- Goals per team

The following plot depicts the number of goals scored by team and by season.

* Boca Juniors, Racing Club and River Plate seem to be the teams that score the most, throughout the seasons analyzed.
* There are teams missing in some seasons due to relegation (which, as the number of matches per season, constantly changed in this league during this period)


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#-------------------
#   Data processing
#-------------------
golesDeLocal = partidos %>%
    group_by(temporada, local) %>%
    dplyr::summarise(goles = sum( local_goles))
colnames(golesDeLocal)[which(names(golesDeLocal) == 'local')] <- 'equipo'

golesDeVisitante = partidos %>%
    group_by(temporada, visitante) %>%
    dplyr::summarise(goles = sum(visitante_goles))
colnames(golesDeVisitante)[which(names(golesDeVisitante) == 'visitante')] <- 'equipo'

goles_x_equipo =  merge( x = golesDeLocal, y = golesDeVisitante, by=c("temporada", "equipo"))
goles_x_equipo$nGoles = goles_x_equipo$goles.x + goles_x_equipo$goles.y 
```

```{r, echo=FALSE}
#----------------
#     PLOT
#----------------
GOLES.X.EQ  = ggplot(data = goles_x_equipo, aes(x=equipo, y=nGoles, fill=nGoles, color='black')) +
            geom_bar(stat='identity') +
            facet_grid(temporada ~ . ) +
            fuentes_customizadas+        
            theme(axis.text.x = element_text(angle = 55, hjust = 1))+
            ylab("Goals")+
            xlab("Team")+
            scale_fill_viridis(begin=0,end=1,direction=-1,discrete=FALSE,option='magma') +
            ggtitle("Total goals scored by teams per season") +
            theme(strip.text.y = element_text(angle=0)) 
```

```{r, echo=FALSE, fig.align = 'center', fig.width=10, fig.height=8}
GOLES.X.EQ  
```



## 2.- Cards per team 

Are there teams that are booked more than others? 

* For yellow cards: the most intensive seasons seem to be 2016-17 and 2022 overall
* For the former, several teams have accumulated a high number of yellow cards, while for the latter Boca Juniors seems to have the excessive amount fo cards
* For red ones, Rosario Central appears to be the most aggressive one during the 2016-17 season while San Lorenzo repeats that trend during the 2021.

```{r, warning=FALSE,message=FALSE, echo=FALSE}
#-------------------
#   Data processing
#-------------------

tarjetas_local = partidos %>%  group_by(temporada, local) %>% 
    dplyr::summarise(amarillas = sum(local_amarillas),
              rojas = sum(local_rojas))
colnames(tarjetas_local)[which(names(tarjetas_local) == 'local')] <- 'equipo'

tarjetas_visitante = partidos %>%  group_by(temporada, visitante) %>% 
    dplyr::summarise(amarillas = sum(visitante_amarillas),
              rojas = sum(visitante_rojas))
colnames(tarjetas_visitante)[which(names(tarjetas_visitante) == 'visitante')] <- 'equipo'

tarj = merge(x=tarjetas_local, y=tarjetas_visitante, by=c('temporada', 'equipo'))
tarj$equipo = as.factor(tarj$equipo)
tarj$amarillas = tarj$amarillas.x + tarj$amarillas.y
tarj$rojas = tarj$rojas.x + tarj$rojas.y
```

```{r, echo=FALSE}
#----------------
#     PLOT
#----------------
AMARILLAS.X.EQ = ggplot(data = tarj, 
                    aes(x=equipo, y=amarillas, fill=amarillas))+
            geom_bar(stat='identity') +
            facet_grid(temporada ~ . ) +
            fuentes_customizadas+ 
            theme(axis.text.x = element_text(angle = 55, hjust = 1))+
            ylab("Number of yellow cards")+
            xlab("Team")+
            ggtitle("Number of yellow cards per team")+
            theme(strip.text.y = element_text(angle=0)) +
            scale_fill_viridis(begin=0,end=1,direction=-1,discrete=FALSE,option='viridis')

ROJAS.X.EQ = ggplot(data = tarj, 
                    aes(x=equipo, y=rojas, fill=rojas))+
            geom_bar(stat='identity') +
            facet_grid(temporada ~ . ) +
            fuentes_customizadas+ 
            theme(axis.text.x = element_text(angle = 55, hjust = 1))+
            ylab("Number of red cards")+
            xlab("Team")+
            ggtitle("Number of red cards per team")+
            theme(strip.text.y = element_text(angle=0)) +
            scale_fill_viridis(begin=0,end=1,direction=-1,discrete=FALSE,option='viridis')
```


```{r, echo=FALSE, fig.align = 'center', fig.width=10, fig.height=8}
AMARILLAS.X.EQ
```

```{r, echo=FALSE, fig.align = 'center', fig.width=10, fig.height=8}
ROJAS.X.EQ
```

## 3.-  Referee's cards performance

Are there more/less strict referees in terms of bookings? Not all of them delivered justice in every season, but some of them appear to be more prone to showing yellow and red cards.


```{r, warning=FALSE,message=FALSE, echo=FALSE}
#-------------------
#   Data processing
#-------------------
arbitros = partidos %>% group_by(temporada, arbitro) %>%  
    dplyr::summarise(ama_al_local = sum(local_amarillas),
              ama_al_visitante = sum(visitante_amarillas),
              roja_al_local = sum(local_rojas),
              roja_al_visitante = sum(visitante_rojas))

arbitros$amarillas_totales = arbitros$ama_al_local + arbitros$ama_al_visitante
arbitros$rojas_totales = arbitros$roja_al_local + arbitros$roja_al_visitante

ama_df <-arbitros %>% arrange(amarillas_totales)

roja_df <- arbitros %>%  arrange(rojas_totales)
```

```{r, echo=FALSE}
#----------------
#     PLOT
#----------------
ARBITROS_AMA  <- ggplot(data = ama_df,
                       aes(x =arbitro, y = amarillas_totales, fill =amarillas_totales)) +
    geom_bar(stat = "identity", width = 0.75) +
    coord_flip() + fuentes_customizadas+ 
    scale_fill_viridis(begin=0,end=1,direction=-1,discrete=FALSE,option='viridis')+
    labs(x = "Season", y = "Number of yellow cards", title = "") +
    facet_grid(.~temporada)+
    ggtitle("Yellow cards shown by referee per season")

ARBITROS_ROJA <- ggplot(data = roja_df,
                       aes(x =arbitro, y = rojas_totales, fill = rojas_totales)) +
    geom_bar(stat = "identity", width = 0.75) +
    coord_flip() + fuentes_customizadas+ 
    scale_fill_viridis(begin=0,end=1,direction=-1,discrete=FALSE,option='viridis')+
    labs(x = "Season", y = "Number of red cards", title = "") +
    facet_grid(.~temporada) +
    ggtitle("Red cards shown by referee per season")
```

```{r, echo=FALSE, fig.align = 'center', fig.width=10, fig.height=8}
ARBITROS_AMA
```

```{r, echo=FALSE, fig.align = 'center', fig.width=10, fig.height=8}
ARBITROS_ROJA
```


# PREDICTING RESULTS WITH STATISTICAL MODELLING

We're going to use the last season available for the most updated data (i.e. season 2022). 


## 1.- Season development and winner

```{r, warning=FALSE, message=FALSE, echo=FALSE}
#-------------------
#   Data processing
#-------------------
data = partidos %>% filter(temporada == '2022')
data = data[c('local', 'visitante', 'semana', 'resultado', 'local_goles', 'visitante_goles')]

season_development = data.frame()

for (i in unique(data$local)) {
        result = data %>% filter(local == i | visitante == i) #filtro x equipo
        result = result[c('local', 'visitante' ,'resultado', 'semana')]
        result$equipo = i
        result$puntos = 0
        #Generar una columna con case when
        df_temp= result %>%
          mutate(puntos = case_when(
            (local==i & resultado == 'empate') ~ 1,
            (local==i & resultado == 'local') ~ 3,
            (local==i & resultado == 'visitante') ~ 0,
            (visitante==i & resultado == 'empate') ~ 1,
            (visitante==i & resultado == 'local') ~ 0,
            (visitante==i & resultado == 'visitante') ~ 3
            )) 
        df_temp$puntos_acum = cumsum(df_temp$puntos)
        #df_temp = df_temp[c('semana' ,'equipo','puntos')]
        season_development = rbind(season_development, df_temp)
}
        
GANADOR = season_development[which.max(season_development$puntos_acum),]$equipo

#-------------------
#   PLOT
#-------------------

EVOL.TEMPORADA =
ggplot(data = season_development) + 
                geom_line(aes(x=semana, y=puntos_acum, 
                          color=equipo
                          ), alpha=0.7) +
                geom_point(aes(x=semana, y=puntos_acum, 
                           color=equipo))+
    theme_bw() + 
    fuentes_customizadas+ 
    xlab("Matchweek")+
    ylab("Accumulated points")+
    ggtitle("Development of the season")+
    theme(legend.position = 'bottom')+
    labs(color='Teams') +
    geom_point(aes(x=max(semana),y=max(puntos_acum)), fill='grey',colour="black",
                   shape=21, size=2.5) +
    annotate(geom='text', 
             x=max(season_development$semana)-2, 
             y=max(season_development$puntos_acum) +2, label=GANADOR,color="black")
```

```{r, echo=FALSE, fig.align = 'center', fig.width=10, fig.height=5}

EVOL.TEMPORADA

```

First, let's look at the development of the season, whose winner was `r  GANADOR`. 

Now to perform predictions, we'll apply the most basic distribution for discrete numbers (in our case, this will be the number of goals), that is, the Poisson distribution.

Such distribution assumes:

* The probability of an event occurring in a given time interval does not vary with time
* The events occur at random
* The events occur independently

This distribution has only one parameter, lambda, which is the average number events per period of time. In terms of our study case, lambda will represent the average number of goals scored per 90-min match.


```{r, warning=FALSE, message=FALSE, error=TRUE, echo=FALSE}

HOME.AVG = mean(data$local_goles) 
AWAY.AVG = mean(data$visitante_goles)
```

We can directly calculate the average number of goals in the desired season, so that the home team scores in average `r round(HOME.AVG,2)` while the away team scores `r round(AWAY.AVG, 2)`


## 2.- Number of goals per match (observed vs. expected)

The plot below shows the proportion of matches where goals were scored by the home and the away team (observed, bars), while also showing the estimated by the corresponding Poisson distributions using thee average goals as before (estimated, lines).

* From the observed distributions, the away team seem to have a much higher proportion of matches without scoring any goals, and that number is greater than for the home team.
* For scoring one goal, the proportion is similar, while the trend shows a much higher proportion of matches where the home team scored 2 to 5 goals in comparison with the away team. 
* The fitted model does not seem to deviate greatly from the actual seen numbers. 


```{r, warning=FALSE, message=FALSE, error=TRUE, echo=FALSE}
#-------------------
#   Data processing
#-------------------
data_temp = reshape2::melt(data, direction = 'long',
                          value.name = 'equipo',
                          variable.name = "condicion",
                          id.vars=c("semana", "local_goles", 'visitante_goles')) 

data_temp_local = data_temp %>% filter(condicion =='local')
data_temp_local = data_temp_local[,c("semana", 'local_goles','condicion','equipo' )]
colnames(data_temp_local)[which(names(data_temp_local) == 'local_goles')] <- 'n_goles'

data_temp_vis = data_temp %>% filter(condicion =='visitante')
data_temp_vis = data_temp_vis[,c("semana", 'visitante_goles','condicion','equipo' )]
colnames(data_temp_vis)[which(names(data_temp_vis) == 'visitante_goles')] <- 'n_goles'

data_long = rbind(data_temp_local, data_temp_vis)

prop_obs_local = data_long %>% 
    filter(condicion=='local') %>% 
    group_by(n_goles) %>% 
   dplyr::summarise(prop_goles=n()/nrow(.))  
prop_obs_local$observacion = 'observada'
prop_obs_local$condicion = 'local'

prop_obs_vis = data_long %>% 
    filter(condicion=='visitante') %>% 
    group_by(n_goles) %>% 
    dplyr::summarise(prop_goles=n()/nrow(.))  
prop_obs_vis$observacion = 'observada'
prop_obs_vis$condicion = 'visitante'

dpois_obs = rbind(prop_obs_local, prop_obs_vis)

prop_est_local = data.frame(
    prop_goles = dpois(0:max(data_long$n_goles),mean(data$local_goles)),
    n_goles = seq(0,max(data_long$n_goles, by=1)))
prop_est_local$observacion = 'estimada'
prop_est_local$condicion = 'local'

prop_est_vis = data.frame(
    prop_goles = dpois(0:max(data_long$n_goles), mean(data$visitante_goles)),
    n_goles = seq(0,max(data_long$n_goles, by=1)))
prop_est_vis$observacion = 'estimada'
prop_est_vis$condicion = 'visitante'

dpois_est = rbind(prop_est_local, prop_est_vis)

specific.val = ppois(q=1, lambda= mean(data$local_goles), lower.tail=FALSE)

```


```{r, warning=FALSE, message=FALSE, echo=FALSE}
#----------------
#     PLOT
#----------------
DISTRIB.POI = 
    ggplot() +
    geom_bar(data = dpois_obs, 
             aes(x = n_goles, y = prop_goles, fill=condicion, color=condicion),
             stat = "identity", position='dodge',  width=0.8) +
        scale_fill_manual(values= c('local'= "springgreen", 
                                    'visitante'= "steelblue"),
                      labels = c("local"= 'Home (obs.)' , 'visitante'='Away (obs.)'),
                      name=""
                      ) +
        fuentes_customizadas+ 
    xlab("Number of goals per match")+
    ylab("Proportion of matches")+
    ggtitle("Observed and expected number of goals scored and\nproportion of matches per team acting as home/away")+
    scale_x_continuous(breaks=0:5, labels=0:5) +
    theme(legend.position = 'bottom')+
    labs(fill = "Condition")


DISTRIB.POI = DISTRIB.POI  + 
    geom_line(data=dpois_est, aes(x=n_goles, y=prop_goles, 
                                 color = condicion), linewidth = 1.4) +
    scale_color_manual(values= c('local'= "red", 
                                    'visitante'= "blue"),
                      labels = c("local"= 'Home (est.)', 
                                 'visitante'='Away (est.)'),
                      name="") +
    geom_point(data = dpois_est, 
               aes(x = n_goles, y = prop_goles), color='black', size=2) +
    labs(fill = "Condition")
```

```{r,echo=FALSE, fig.align = 'center', fig.width=6, fig.height=6}
DISTRIB.POI 
```


We can go a step further with the simple Poisson model, estimating the probability of specific events. For example, what is the probability of the h ome team scoring 2 or more goals? we're asking for $P( \ge 2   |  Home)$, which is `r specific.val`.
 

## 3.- Difference goal result: most probable scenarios

What about draws? for this, we need to know when the difference between the home and away team Poisson distributions is 0. The distribution that results from subtracting one Poisson from another is called a Skellam distribution.


```{r, warning=FALSE, message=FALSE, echo=FALSE}
#-------------------
#   Data processing
#-------------------
data$gol_diff = data$local_goles - data$visitante_goles 

diff_gol = data %>% group_by(gol_diff) %>%
    dplyr::summarize(prop_goles_obs=n()/nrow(.)) %>%
     inner_join(data.frame(gol_diff=-4:5,
                prop_goles_estimado=skellam::dskellam(-4:5,mean(data$local_goles),
                mean(data$visitante_goles))),by=c("gol_diff"))
```

```{r, echo=FALSE}
#----------------
#     PLOT
#----------------
SKELLAM.DISTRIB = 
    ggplot() +
    geom_bar(data = diff_gol, 
             aes(x = gol_diff, y = prop_goles_obs, fill='grey21', color='grey21'),
             stat = "identity",  width=0.8) +
        fuentes_customizadas+
    theme(legend.position='none') + 
    xlab("Number of goals difference (Home - Away team)")+
    ylab("Proportion of matches")+
    ggtitle("Skellam distribution for Home-Away results")
    scale_x_continuous(breaks=-4:5, labels=-4:5)

SKELLAM.DISTRIB =SKELLAM.DISTRIB +
    geom_line(data=diff_gol, aes(x = gol_diff, y = prop_goles_obs)) + 
    geom_point(data=diff_gol, aes(x = gol_diff, y = prop_goles_obs))
```


```{r, echo=FALSE, fig.align = 'center', fig.width=6, fig.height=6}
SKELLAM.DISTRIB 
```


With this plot we can see that for the Argentinean League, the most probable scenario is a 0-0 draw. Then, there is a higher probability of the home team winning by one goal and in third place, there is almost the same probability of the home team winning by two goals and the away team, winning by one goal.


## 4.- Build a GLM odel

Finally, lets build a GLM model with which we can make some specific analysis on the participating teams. 

```{r, warning=FALSE, message=FALSE, error=TRUE, echo=FALSE}
#-------------------
#   Data processing
#-------------------
df_model =  rbind(
            data.frame(n_goles = partidos$local_goles,
                       equipo = partidos$local,
                       oponente = partidos$visitante,
                       localia = 1),
            data.frame(n_goles = partidos$visitante_goles,
                       equipo = partidos$visitante,
                       oponente = partidos$local,
                       localia = 0))
 
poi_model = glm(n_goles ~ equipo + oponente + localia,
                family=poisson(link=log),data=df_model)

```

```{r, echo=FALSE}
nPartidos = partidos %>%
    filter(temporada=='2022') %>%
    dplyr::summarise(nPartidos = max(semana))

```

Note that if we apply this model to a small sample size (lets say when the season is going through the 8th matchweek, where each team in the season would've played `r nPartidos` matches), the accuracy of this approximation can vary significantly.


```{r, message=FALSE, warning=FALSE, echo=FALSE} 
pander::pander(summary(poi_model))
```

To make our analysis, bear in mind that the results are displayed in the log scale, so we have to calculate $e ^ {estimate_i}$. Does the home factor weights? yes! since $e^{0.25}  = 1.28$. 

Then we can analyze the performance of each team, using the "equipo" values. Positive values mean that the team scores more than average and negative, the opposite. For example:

* For River Plate: $e^{0.78} = 2.18$ [River Plate scores 2.18 goals more than the average team]
* For Quilmes: $e^{-0.3} = 0.74$ [Quilmes scores 0.74 goals than the average team]

Finally, the "oponente" values penalize/reward teams based on the quality of their opposition. This mimics the defensive strength of each team"

* For San Lorenzo, $e^{-0.26} = 0.77$
* For River plate, $e^{-0.5} = 0.61$
In other words, if you're playing these teams, you're less likely to score against River plate than San Lorenzo.

If we plot the "equipo" and the "oponente" estimates per team, separatedly, we can get a peek of which team has the most probability of score above/below average, and which is the most/least defensive (in terms of average goals received).

#### 4a. Attacking strenght estimated by the model:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#-------------------
#   Data processing
#-------------------
df=data.frame(summary(poi_model)$coefficients)
df<- cbind(newColName = rownames(df), df)
rownames(df) <- 1:nrow(df)
colnames(df) = c("equipo", 'estimate', 'std_error','z_value','p_value')

potenciales_goles = df %>% dplyr::filter(stringr::str_detect(equipo, 'equipo'))
potenciales_goles$estimate_lineal = exp(potenciales_goles$estimate)
potenciales_goles$equipo <- stringr::str_replace(potenciales_goles$equipo , "equipo", "")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
mentality = data.frame(poi_model$coefficients)
mentality <- cbind(newColName = rownames(mentality ), mentality )
rownames(mentality ) <- 1:nrow(mentality )
colnames(mentality ) = c("equipo", 'estimate')

most_offensive = mentality[which.max(mentality$estimate),]$equipo
most_offensive = gsub('equipo','',most_offensive)

most_deffensive = mentality[which.min(mentality$estimate),]$equipo
most_deffensive = gsub('oponente','',most_deffensive)

```


```{r, echo=FALSE}
#----------------
#     PLOT
#----------------
POTENCIAL.GOLES =  ggplot(potenciales_goles, 
                          aes(x = reorder(equipo, estimate_lineal),
                              y = estimate_lineal, 
                              color=estimate_lineal,
                              fill=estimate_lineal)) +
                   geom_bar(stat = "identity", width = 0.75) +
    coord_flip() +
    fuentes_customizadas +
    ylab("Goals scored relative to average (GLM Poi model estimates)")+
    xlab("Team")+
    ggtitle("Offensive strenght estimation per team by GLM Poi") +
    labs(fill = "")+
    scale_fill_gradient(low = "blue4", high = "green")
```

Given the model's estimates, the largest estimate for the offensive power is for `r most_offensive`.

```{r, fig.align = 'center', fig.width=8, fig.height=6}
POTENCIAL.GOLES
```


#### 4b.  Deffensive strenght estimated by the model:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#-------------------
#   Data processing
#-------------------
poder_defensivo = df %>% dplyr::filter(stringr::str_detect(equipo, 'oponente'))
poder_defensivo$estimate_lineal = exp(poder_defensivo$estimate)
poder_defensivo$equipo <- stringr::str_replace(poder_defensivo$equipo , "oponente", "")
```

```{r, echo=FALSE}
#----------------
#     PLOT
#----------------
POTENCIA.DEF =  ggplot(poder_defensivo, 
                          aes(x = reorder(equipo, estimate_lineal),
                              y = estimate_lineal, 
                              color=estimate_lineal,
                              fill=estimate_lineal)) +
                   geom_bar(stat = "identity", width = 0.75) +
    coord_flip() +
    fuentes_customizadas +
    ylab("Opponency to receive goals relative to average (GLM Poi model estimates)")+
    xlab("Team")+
    ggtitle("Deffensive strenght estimation per team by GLM Poi") +
    labs(fill = "")+
    scale_fill_gradient(low = "blue4", high = "green")
```

Given the model's estimates, the largest estimate for the defensive power is for `r most_deffensive`.


```{r,  fig.align = 'center', fig.width=8, fig.height=6}
POTENCIA.DEF
```



